---
title: "R Programming module 8"
author: "Maimuna"
format: pdf
editor: visual
---

### 1. Data frames

```{r}
name <- c("Raju", "Raj", "Raba", "Rahul", "Rimi")
age <- c(11, 9, NA, 10, 5)
gender <- c(NA, "boy", "girl", "boy", "girl")
loc <- c(1, 2, 1, 1, NA)

# Creating data frame
df <- data.frame(name, age, sex=gender, loc)

# Convert categorical variables to factor
df$sex <- factor(df$sex)
df$loc <- factor(df$loc, labels = c("Urban", "Rural"))

df
```


### 2. Some useful functions

```{r}
names(df)
dim(df)
summary(df)
summary(df$age)
table(df$sex)
```


### 3. Ordering data frames

```{r}
# by age
df

df[order(df$age), ]
```


### 4. Handling missing data
```{r}
na.omit(df)  # drops all rows containing NA
mean(df$age, na.rm = T)
```

### 5. Adding new column or rows

```{r}
df$place <- c("UK", "BN", "PK", "IN", "BN")
df
```

```{r}
df$loc <- NULL
df
```

```{r}
# rbind for rows
df1 <- data.frame(id = 1:4, height = c(120, 150, 132, 122),
                        weight = c(44, 56, 49, 45))
df1

df2 <- data.frame(id = 5:6, height = c(119, 110), weight = c(39, 35))
df2

rbind(df1, df2)

df3 <- data.frame(location = c("UK", "CZ", "CZ", "UK"))
df3

cbind(df1, df3)
```


### 6 Analyse a subset of data

```{r}
df_girl <- df[df$sex == "girl", ]
df_girl

sum(df_girl$age, na.rm = T)
```

### Exercise 8.1

The data mtcars comprises fuel consumption and 10 aspects of automobile design
and performance for 32 automobiles. Load the data by running data(mtcars)

1. Obtain the variable list of the data frame mtcars

```{r}

```


2. How many observations and variables do the mtcars data have?

```{r}

```


3. Check the types of the variables of mtcars

```{r}

```


4. Rename the variable hp to horsepower

```{r}

```


5. Order the dataset in ascending order of the variable mpg (miles per gallon)

```{r}

```


6. Convert the variable cyl (number of cylinders) to factor type variable

```{r}

```


7. Create a subset of the mtcars dataset where mpg is less than 30, retaining only the first five variables. Save the resulting dataset as mtcars_subset.

```{r}

```


8. Find the mean, median, mode, range, standard deviation, and IQR of the variable Miles/(US) gallon (mpg)

```{r}

```


9. Find the frequency table of Number of cylinders (cyl)

```{r}

```


10. Find the 2-way contingency table of Number of cylinders (cyl) and Number of forward gears (gear)

```{r}

```


### 8 Frequency table
```{r}
table(mtcars$mpg)
```

```{r}
mtcars$mpg_cat <- cut(mtcars$mpg, breaks = c(10, 20, 30, 40),
                      labels= c("low", "med", "high"), right = T)
mtcars$mpg_cat

table(mtcars$mpg_cat)
table(mtcars$mpg_cat, mtcars$cyl)
```

