---
title: "R Programming module 7"
author: "Maimuna"
format: pdf
editor: visual
---

### 1. Control Flow - Choices (if, if else, case)

```{r}
x <- sample(1:10, size = 1)
x

if(x >= 5) {
  "x is greater than or equal to 5"
} else {
  "x is smaller than 5"
}
```

```{r}
# grading system

grade <- function(x) {
  if(x >= 90) y = "A"
  else if(x >= 80 & x < 90) y = "B"
  else if(x >= 60 & x < 80) y = "C"
  else y = "D"
  
  return(y)
}

grade(90)
```


```{r}
library(dplyr)

grade_case <- function(x) {
  
  y <- case_when(x >= 90 ~ "A",
                 x >= 80 & x < 90 ~ "B",
                 x >= 60 & x < 80 ~ "C",
                 x < 60 ~ "D")
  return(y)               
}

grade_case(67)
```


```{r}
x <- 1:10
x

dplyr::if_else(x %% 2 == 0, "even", "odd")
```

### 2. Loops

```{r}
v <- c("this", "is", "a", "for", "loop")

for(i in v) {
  print(i)
}
```

```{r}
for(i in 1:10) {
  if(i < 3) next
  if(i >= 8) break
  print(i)
}
```

```{r}
z <- 1
while(z < 4) {
  z = z+1
  print(z)
}
```

### Exercise 7.1

1. Write a function max_of_three(a, b, c) that returns the largest of three
numbers using only if and else. Do not use max(), sort(), or which.max().

```{r}
max_of_three <- function(a, b, c) {
  
  if(a > b && a > c) return(a)
  else if(b > a && b > c) return(b)
  else return(c)
}

max_of_three(1, 10, 5)
```


2. Write a function piecewise(x) that returns ...
   Use an if/else if/else structure.
   
```{r}
piecewise <- function(x) {
  
  if(x < 0) {
    return((x * x) + 1)
  } else if(x >= 0 && x < 5) {
    return(2*x)
  } else {
    return(x^(0.5))
  }
}

piecewise(-10)
piecewise(3)
piecewise(7)
```
   

3. Write a function valid_date(day, month) that returns TRUE when the supplied
day is valid for the supplied month and FALSE otherwise. Assume a non-leap year
and that month is an integer from 1 to 12.

```{r}
valid_date <- function(day, month) {
  
  if(month == 1 | month == 3 | month == 5 | month == 7 | month == 8 | 
     month == 10 | month == 12) {
    if(day >= 1 && day <= 31) {
      return(T)
    } else {
      return(F)
    }
  } else if(month == 4 | month == 6 | month == 9 | month == 11) {
    if(day >= 1 && day <= 30) {
      return(T)
    } else {
      return(F)
    }
  } else if(month == 2) {
    if(day >= 1 && day <= 28) {
      return(T)
    } else {
      return(F)
    }
  } else {
    return(F)
  }
}

valid_date(31, -1)
valid_date(1, 10)
```


4. Write a function is_prime(n) that uses a for loop to return TRUE if n is
prime and FALSE otherwise. Return FALSE for n < 2.

```{r}
is_prime <- function(n) {
  
  if(n < 2) return(F)
  if(n == 2) return(T)

  for(i in 2:(n-1)) {
    if(n %% i == 0) {
      return(F)
    }
  }
  
  return(T)
}

is_prime(11)
```


5. Using a for loop, create a vector containing the first 15 Fibonacci numbers.
Start with 0 and 1.

```{r}
fibonacci <- c(0, 1)

for(i in 3:15) {
  fibonacci[i] <- fibonacci[i-1] + fibonacci[i-2]
}

print(fibonacci)
```


6. Write a function first_repeated_value(x) that uses a for loop to return the
first value that appears for a second time while scanning x from left to right.
Return NA if there is no repeated value. Do not use duplicated() or unique().

```{r}
first_repeated_value <- function(x) {
  
  if(length(x) < 2) return(NA)
  
  seen <- c()
  
  for(item in x) {
    if(item %in% seen) return(item)
    
    seen <- c(seen, item)
  }
  
  return(NA)
}

first_repeated_value(x = c(2, 3, 5, 3, 3))
```


7. Write a function find_gcd(a, b) that uses a while loop to find the greatest
common divisor of two positive integers using the Euclidean algorithm. Do not
use a pre-written GCD function.

```{r}
gcd <- function(a, b) {
  
  while(b != 0) {
    remainder <- a %% b
    a <- b
    b <- remainder
  }
  
  return(a)
}

gcd(12, 4)
```


### 3. Base functionals

```{r}
mat <- matrix(1:15, nrow = 3)
mat

apply(mat, 1, sum)   # 1 for row
apply(mat, 2, sum)   # 2 for col

rsum <- rep(NA, nrow(mat))
for (i in 1:nrow(mat)) {
  rsum[i] <- sum(mat[i, ])
}
rsum
```

lapply() returns a list after applying a function on each element of an R object

sapply() is similar to lapply() but it returns a atomic vector instead of a list

```{r}
lres <- list(x = 1:3, y = matrix(1:4, 2))
lres

lapply(lres, sum)
sapply(lres, sum)
```

