---
title: "R Programming module 6"
author: "Maimuna"
format: pdf
editor: visual
---

### 1. Function
```{r}
right_triangle_area <- function(base, height) {
  area <- 0.5 * base * height
  return(area)
}

right_triangle_area(2, 3)
```

### 2. Simulating rolling dice

```{r}
roll <- function() {
  dice <- 1:6
  out <- sample(x = dice, size = 1)
  return(out)
}

roll()
roll
```

### 3. Functions with arguments

```{r}
# toss a fair coin n times, number of heads = ?
# SRSWR

coin_head <- function(n) {
  head <- sample(c(0, 1), size = n, replace = T)
  total_head <- sum(head)
  return(total_head)
}

coin_head(1000)

```


### Exercise 6.1

1. Write an R function roll_dice() that rolls a dice n times (where n is 1 or more) and returns the sum of the dice rolls.

```{r}
roll_dice <- function(n) {
  
  if (n < 1) {
    stop("n must be 1 or more times.")
  }
  
  dice <- sample(1:6, size = n, replace = T)
  sum(dice)
}

roll_dice(10)
```


2. Write an R function toss_bias_coins() that tosses a biased coin n times (where n is 1 or more times) and returns the number of heads. The probability of getting heads is 0.70.

```{r}
toss_bias_coins <- function(n) {
  
  if(n < 1) {
    stop("n must be greater than or equal to 1")
  }
  
  tosses <- rbinom(n = n, size = 1, prob = 0.7)
  return(sum(tosses))
}

toss_bias_coins(10)
```


3. Write an R function that can simulate flipping a fair or biased coin. The probability of getting heads should not always be 0.70, but any value between 0 and 1!

```{r}
toss_bias_coins_2 <- function(n, probability) {
  
  if(n < 1) {
    stop("n must be greater than or equal to 1")
  }
  
  if(probability < 0 | probability > 1) {
    stop("probability range : [0,1]")
  }
  
  tosses <- rbinom(n = n, size = 1, prob = probability)
  return(sum(tosses))
}

toss_bias_coins_2(10, 0.8)
```

