---
title: "R Programming module 4"
author: "Maimuna"
format: pdf
editor: visual
---

### 1. Matrics

```{r}
# creating matrix using matrix()

A = matrix(
  1:9,
  nrow = 3,
  ncol = 3,
  byrow = T     # default: byrow = FALSE
)
A
```


```{r}
# Creating matrix using rbind() or cbind()

B <- rbind(1:3, 11:13, 111:113)
C <- cbind(letters[1:2], c("k", "l"), letters[13:14])


B
cat("\n")
C
```

```{r}
# Creating matrix by adding dim attribute

z <- 101:108
dim(z) <- c(4, 2)
z
```


### 2. Arrays
```{r}
# Creating array using array()

D <- array(1:8, dim = c(2, 2, 2))
D
```


```{r}
# Creating array by adding dim attribute
z <- 1:18
dim(z) <- c(2, 3, 3)
z
```


### 3. Calculations on matrices

```{r}
meow_mat <- matrix(1:16, nrow = 4, ncol = 4, byrow = T)
rownames(meow_mat) <- c("A", "B", "C", "D")
colnames(meow_mat) <- c("a", "b", "c", "d") 

meow_mat
```


```{r}
mat1 <- matrix(c(2, 0, 1, 1), nrow = 2)
mat2 <- matrix(c(1, 1, 0, 2), nrow = 2)
mat1
mat2

mat_add <- mat1 + mat2
mat_add

mat_element_mult <- mat1 * mat2
mat_element_mult

mat_mult <- mat1 %*% mat2
mat_mult

```


### 4. Basic martrix functions

```{r}
# transpose
meow_mat_t <- t(meow_mat)
meow_mat_t

# diagonal element
diag(meow_mat)

ncol(meow_mat)
nrow(meow_mat)
colSums(meow_mat)
rowSums(meow_mat)
```


### Exercise 4

1. Create a vector called x consisting of the first fifteen integers of the number line.

2. Use the function dim() to assign dimension to vector x with three rows and five columns. What is the class of x now?

```{r}
# 1.

x_ex4 <- 1:15
x_ex4

dim(x_ex4) <- c(3, 5)
x_ex4
```
   

3. Given the following matrices ...

   i. Calculate A^(T)b
   ii. Find the inverse of matrix A
   iii. Solve the equation for x, where Ax = b
   
```{r}
A_ex4 <- matrix(c(2, 0, 7, 7, 9, 4, 5, 8, 0, 1, 5, 7, 0, 4, 1, 4), nrow = 4)
A_ex4

b_ex4 <- matrix(c(-1, 6, 0, 9), nrow = 4)
b_ex4

# i
AT_b <- t(A_ex4) %*% b_ex4
AT_b

# ii
solve(A_ex4)

# iii
solve(A_ex4, b_ex4)

```
   

4. i. Generate a vector x0 of order 20 with all elements as 1

ii. Generate a vector x1 of order 20 with elements randomly selected from 30:70, consider a seed 80

iii. Create a matrix X with the first column x0 and the second column x1

iv. Generate a vector Y of order 20 using the equation yi = 1.2 + 1.8x1 + epsilon(i), where epsilon(i) ~ N(0,9) 

v. Obtain the value of (X'X)^(-1)X'Y

```{r}
# i
x0 <- rep(1, times = 20)
x0

# ii
set.seed(80)
x1 <- sample(30:70, 20, replace = T)
x1

# iii
X <- matrix(c(x0, x1), ncol = 2)
X

# iv
epsilon <- rnorm(20, mean = 0, sd = 9)
Y <- 1.2 + (1.8 * x1) + epsilon
Y

# v
value <- (solve(t(X) %*% X)) %*% (t(X)) %*% Y
value
```


### 5. Factors

Factors look like strings, but behave like integers

```{r}
# Creating factor using factor()

fc = factor(x = c(1, 2, 2, 3, 4, 4, 3, 1, 4, 2, 2, 3, 1, 4, 1),
             levels = c(1, 2, 3, 4),
             labels = c("east", "west", "north", "south"))

fc

typeof(fc)
as.character(fc)
```

```{r}
# not providing levels
fac1 = factor(c("Male", "Female", "Male",
               "Male", "Female", "Male", "Female"))
fac1

typeof(fac1)
attributes(fac1)
```

```{r}
# providing levels 
fac2 = factor(c("Male", "Female", "Male",
               "Male", "Female", "Male", "Female"), 
             levels = c("Male", "Female"))
fac2
```

### 6. Lists

List is a vector with heterogeneous elements, i.e., each element of a list can be any type

```{r}
ls1 <- list(1:3,
            "m",
            c(T, F, F, T, T),
            c(2.5, 1.2, 0.8, 5.5))

ls1

typeof(ls1)
is.list(ls1)
str(ls1)
```

```{r}
# Lists are sometimes called recursive vectors
l3 <- list(list(list("First list")))
l3
str(l3)
```


### 7. Data Frame

```{r}
name <- c("cat", "dog", "duck", "cow")
language <- c("meow", "brrr", "quack", "moo")
age <- c(2, 5, 1, 3)

df <- data.frame(name, language, age)
df

typeof(df)
is.data.frame(df)
attributes(df)
str(df)
names(df)
summary(df)
head(df)
tail(df)
```

### 8. Tibble

```{r}
# Load the tibble package
library(tibble)

my_data <- tibble(
  name = c("Samir", "Amir", "Aman"),
  age = c(25, 30, 35),
  city = c("Dhaka", "Khulna", "Jashore")
)
my_data
```

```{r}
# mtcars data set as tibble

mtcars_t <- as_tibble(mtcars)
mtcars_t
```

