---
title: "Basic Measures of Demography"
subtitle: "232 Module 1 Pratice"
format: html
editor: visual
---

### The data frame

```{r}
# Pabna District, mid year 2022 (illustrative figures)
age_group <- c("0-4",   "5-9",   "10-14", "15-19", "20-24",
                "25-29", "30-34", "35-39", "40-44", "45-49",
                "50-54", "55-59", "60-64", "65-69", "70-74",
                "75-79", "80+")

male   <- c(124500, 138200, 142300, 131400, 118200,
            104500, 96300,  88700,  78400,  67100,
            56800,  47200,  38500,  28700,  19600,
            11400,  8200)

female <- c(119800, 132400, 136700, 128900, 122600,
            110800, 102400, 94200,  82500,  70200,
            58900,  48400,  39200,  29800,  21200,
            13500,  11800)

pop <- data.frame(age_group, male, female)
pop
```

## Masculinity Proportion (MP)

```{r}
M <- sum(pop$male)
F <- sum(pop$female)
tot_pop <- M + F

MP <- M / tot_pop  * 100
round(MP, 2)
```

In Pabna 2022, roughly 49.56% of the population is male. the remaining 50.44% is female. The slight female majority is typical in Bangladesh and reflects female longevity in adult and older age groups.

## Sex Ratio (SR)

```{r}
SR <- M / F * 100
round(SR, 2)
```

A Sex Ratio of 98.24 means there are 98 males for every 100 females in Pabna 2022. Values below 100 indicate a female majority, values above 100 indicate a male majority.

## Excess or Deficit of Males

```{r}
excess <- M - F
round(excess, 2)

# excess with a custom target
sr_target <- 105
excess_target <- M - (sr_target / 100) * F
round(excess_target, 2)
```

A value of -23300 means Pabna 2022 has a deficit of 23300 males relative to a perfectly balanced population.A positive value would indicate an excess.

A value of -89465 means Pabna 2022 has a deficit of 89465 males relative to a sex ratio of 105.

## Aged Child Ratio (ACR)

```{r}
pop$total <- pop$male + pop$female
p_0_14 <- sum(pop$total[1:3])
p_15_64 <- sum(pop$total[4:13])
p_65 <- sum(pop$total[14:17])

ACR <- p_65 / p_0_14 * 100
round(ACR, 2)
```

There are about 18 aged people (65+ age) against 100 children (aged 0-14) in Pabna 2022. A low value points to a young population; a high value points to an ageing one.

## Dependency Ratios

### Total Dependency Ratio (TDR)

```{r}
TDR <- (p_0_14 + p_65) / p_15_64 * 100
round(TDR, 2)
```

### Child Dependency Ratio (CDR)

```{r}
CDR <- p_0_14 / p_15_64 * 100
round(CDR, 2)
```

### Old Age Dependency Ratio (OADR)

```{r}
OADR <- p_65 / p_15_64 * 100
round(OADR, 2)
```

about 56 people (child and aged) are depended per 100 working people (aged 15-64). Among the 56 depended people, about 47 people are children aged 0-14 and about 9 people are old aged people aged 65-65+.

This is the classic profile of a country still mid way through its demographic transition: most of the dependency burden still sits on the child side.

## Population Pyramid

```{r}
# 1. Male bars, negative so they go to the left of 0
barplot(-pop$male,
        horiz     = TRUE,
        names.arg = pop$age_group,
        space     = 0,
        col       = "#2563eb",
        xlim      = c(-160000, 160000),
        xlab      = "Population",
        main      = "Pabna District, 2022",
        las       = 1,
        axes      = FALSE)

# 2. Female bars, positive so they go to the right; overlay same panel
barplot(pop$female,
        horiz = TRUE,
        space = 0,
        col   = "#f472b6",
        add   = TRUE,
        axes  = FALSE)

# 3. Draw a clean x axis with positive labels on both sides
ticks <- seq(-150000, 150000, by = 50000)
axis(1, at = ticks, labels = abs(ticks))

# 4. Legend
legend("topright",
       legend = c("Male", "Female"),
       fill   = c("#2563eb", "#f472b6"),
       bty    = "n")
```

A wide base with a narrow top means a young, growing population. A narrow base with a wide top means an ageing population. Sudden indentations point to past events such as famines, wars, or migration shocks. Asymmetric bars at older ages reflect female longevity.
