---
title: "Basic Measures of Demography (exercise)"
subtitle: "231 Module 1 Exercise"
author: "Maimuna Rahman"
format: html
editor: visual
---

## Sylhet District 2022 Dataset

```{r}
age_group <- c("0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44", "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80+")

male <- c(196500, 213400, 218700, 198500, 168300, 142800, 128400, 114700, 98300, 82600, 67800, 54300, 42800, 31200, 21400, 12700, 8900)

female <- c(188200, 204800, 209300, 192400, 178500, 152400, 138600, 121200, 102800, 85400, 70200, 56800, 44500, 32800, 23500, 14800, 12200)

syl <- data.frame(age_group, male, female)
syl
```

## 1.1 Masculinity Proportion

```{r}
M <- sum(syl$male)
F <- sum(syl$female)
total_pop <- M + F

MP <- M / total_pop * 100
round(MP, 2)
```

There are about 49.63% of the population is male. The other 50.37% is female in Sylhet 2022.

## 1.2 Sex Ratio

```{r}
SR <- M / F * 100
round(SR, 2)
```

There are about 99 males for every 100 females in Sylhet 2022.

## 2 Aged Child Ratio

```{r}
syl$tot <- syl$male + syl$female
p_0_14 <- sum(syl$tot[1:3])
p_15_64 <- sum(syl$tot[4:13])
p_65 <- sum(syl$tot[14:17])

ACR <- p_65 / p_0_14 * 100
round(ACR, 2)
```

ACR = 12.8 means there are about 13 aged people (aged 65-65+) per 100 child (aged 0-14) in Sylhet 2022.

## 3. Total female population aged 20 to 39

```{r}
tot_F_20_39 <- sum(syl$female[5:8])
round(tot_F_20_39, 2)
```

There are total 590700 females aged from 20 to 39 in Sylhet 2022.

## 4. All three dependency ratios on Sylhet

### 4.1 Total Dependency Ratio

```{r}
TDR <- (p_0_14 + p_65) / p_15_64 * 100
round(TDR, 2)
```

### 4.2 Child Dependency Ratio

```{r}
CDR <- p_0_14 / p_15_64 * 100
round(CDR, 2)
```

### 4.3 Old Age Dependency Ratio

```{r}
OADR <- p_65 / p_15_64 * 100
round(OADR, 2)
```

There are about 62 total people depended (child and old aged) on the working aged people. Among 62 people, about 53 people are children (aged 0-14) and about 7 people are old aged (aged 65-65+).
