---
title: "Fertility Measures Exercise"
subtitle: "232 Module 2 practice"
author: "Maimuna"
format: html
editor: visual
---

## 1. Build the Rajshahi data frame, then compute CBR

```{r}
age_group <- c("15-19", "20-24", "25-29", "30-34", "35-39", "40-44", "45-49")
births <- c(7500, 15000, 11000, 6300, 1900, 510, 75)
female_pop <- c(125000, 115000, 110000, 105000, 95000, 85000, 75000)
L5x <- c(494000, 491000, 488000, 484500, 480000, 474000, 466000)

fert <- data.frame(age_group, births, female_pop, L5x)

total_pop <- 2700000

fert
```

#### Crude Birth Rate (CBR)

```{r}
total_birth <- sum(fert$births)            # 42,285

CBR <- total_birth/ total_pop * 1000
CBR
```

## 2. Annualised Monthly Birth Rate for March 2024

```{r}
march_birth <- 3580
march_day <- 31
march_pop <- 2700000

CBR <- (365/march_day) * march_birth / march_pop * 1000
round(CBR, 2)

```

CBR was 15.66, so March is right at the yearly average pace, only a touch below.

## 3. GFR for Rajshahi

```{r}
total_female <- sum(fert$female_pop)

GFR <- total_birth / total_female * 1000
round(GFR, 2)
```

About 59.56 live births per 1,000 women aged 15 to 49 in Rajshahi 2024. Lower than Khulna's 69.13, pointing to a slightly lower fertility regime.

## 4. Full ASFR vector for Rajshahi, plus a line plot

```{r}
fert$ASFR <- fert$births / fert$female_pop * 1000
fert[, c("age_group", "ASFR")]
```

#### Line plot

```{r}
plot(fert$ASFR,
     main = "ASFR of Rajshahi 2024",
     xlab = "Age Group",
     ylab = "ASFR (Per 1000 women)",
     type = "o",
     xaxt = "n",
     col = "maroon")
axis(1, at = 1:7, labels = fert$age_group)
```

## 5. TFR for Rajshahi

```{r}
n <- 5 # class width
TFR <- n * sum(fert$ASFR) / 1000
round(TFR, 4)
```

Each Rajshahi woman has on average about 1.89 children IN 2024.

## 6. GRR at two different p values

```{r}
GRR_487 <- TFR * 0.487
GRR_485 <- TFR * 0.485

round(c(p_0.487 = GRR_487, p_0.485 = GRR_485), 3)
```

GRR scales linearly with p. A slightly smaller p (fewer daughters per 100 births) gives a slightly smaller GRR.

## 7. NRR for Rajshahi, using L5x from the life table

```{r}
p_fe <- 0.487
l0 <- 100000
NRR <- sum(fert$ASFR * p_fe * fert$L5x / l0) / 1000

round(NRR, 3)
```

GRR was 0.919. NRR is smaller because each age band is weighted by 5Lx / l0, the average woman years lived per radix, which is always less than the full band width n = 5.

whenever any mortality exists. Mortality erodes the reproductive contribution slightly at each age.
